\vspace{1.0cm} \LARGE \bf MERS-CoV spillover at the camel-human interface

### 2017-10-19 mers-structure.tex
Updating AR's affiliations and grants
Fix issue #24

Here, we use an explicit model of metapopulation structure and migration between discrete subpopulations, referred to here as demes \citep{vaughan_efficient_2014}, derived from the structured coalescent \citep{notohara_structured_coalescent_1990}.
Unlike approaches that model host species as a discrete phylogenetic trait of the virus using continuous-time Markov processes (or simpler, parsimony based, approaches) \citep{faria_simultaneously_2013,lycett_h5n8_2016}, population structure models explicitly incorporate distinct sampling patterns, population dynamics within demes and migration between demes.
By estimating independent coalescence rates for MERS-CoV in humans and camels, as well as migration patterns between the two demes, we show that long-term viral evolution of MERS-CoV occurs exclusively in camels.
Our results suggest that spillover events into humans are seasonal and might be associated with the calving season in camels.
Once human MERS-CoV infections are established, however, we find that MERS-CoV is poor at transmitting between humans.
Using Monte Carlo simulations we show that $R_{0}$ for MERS-CoV circulating in humans is much lower than the epidemic threshold of 1.0 and that correspondingly the virus has been introduced into humans hundreds of times.

@@

\section*{Acknowledgements}
We would like to thank Allison Black for useful discussion and advice.
GD is supported by the Mahan postdoctoral fellowship from the Fred Hutchinson Cancer Research Center.
TB is a Pew Biomedical Scholar and is supported by NIH R35 GM119774-01.
AR was supported in part by the European Union Seventh Framework Programme for research, technological development and demonstration under Grant Agreement no. 278433-PREDEMICS and no. 725422-RESERVOIRDOCS, and the Wellcome Trust through project 206298/Z/17/Z.
MAS is partly supported by NSF grant DMS 1264153 and NIH grants R01 HG006139 and R01 AI107034.
We gratefully acknowledge support from NVIDIA Corporation with the donation of parallel computing resources used for this research.

>>>>

Here, we use an explicit model of metapopulation structure and migration between discrete subpopulations, referred to here as demes \citep{vaughan_efficient_2014}, derived from the structured coalescent \citep{notohara_structured_coalescent_1990}.
Unlike approaches that model host species as a discrete phylogenetic trait of the virus using continuous-time Markov processes (or simpler, parsimony based, approaches) \citep{faria_simultaneously_2013,lycett_h5n8_2016}, population structure models explicitly incorporate contrasts in deme effective population sizes and migration between demes.
By estimating independent coalescence rates for MERS-CoV in humans and camels, as well as migration patterns between the two demes, we show that long-term viral evolution of MERS-CoV occurs exclusively in camels.
Our results suggest that spillover events into humans are seasonal and might be associated with the calving season in camels.
Once human MERS-CoV infections are established, however, we find that MERS-CoV is poor at transmitting between humans.
Using Monte Carlo simulations we show that $R_{0}$ for MERS-CoV circulating in humans is much lower than the epidemic threshold of 1.0 and that correspondingly the virus has been introduced into humans hundreds of times.

@@

\section*{Acknowledgements}
We would like to thank Allison Black for useful discussion and advice.
GD is supported by the Mahan postdoctoral fellowship from the Fred Hutchinson Cancer Research Center.
TB is a Pew Biomedical Scholar and is supported by NIH R35 GM119774-01.
AR was supported in part by the European Union Seventh Framework Programme for research, technological development and demonstration under Grant Agreement no. 278433-PREDEMICS and no. 725422-RESERVOIRDOCS, and the Wellcome Trust through project 206298/Z/17/Z.
